---
title: 'Class 6: R Functions'
author: "Anel A15426506"
date: "10/14/2021"
output: pdf_document
---

# A play with Rmarkdown

This is some plain text. I can make things **bold**, I can also make *things italic*.

```{r}
# This is a code chunk
plot(1:10)
```

## R functions

In today's class we are going to write a function together that grades some student work.

Questions for today:

> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: "<https://tinyurl.com/gradeinput>" [3pts]

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```


Let's start with student1 and find their average score.
```{r}
mean(student1)
```

But we want to drop the lowest score...
We could try the **min()** function

```{r}
min(student1)
```

The **which.min()** function looks useful:
```{r}
which.min(student1)
```
This gives the position of the lowest score


```{r}
# This would be the lowest score
student1[ which.min(student1) ]
```
To drop this value I can use minus

```{r}
student1[ -which.min(student1) ]
```

mean student1
```{r}
mean(student1[ -which.min(student1) ])
```

Let's now use mean() to get the average minus the lowest score.

Doesn't work because it gives NA if a student has missing work
```{r}
mean(student2[ -which.min(student2) ])
```

Need to remove the NA elements of the vector
```{r}
# which.min(student2)
mean(student2 [ -which.min(student2) ], na.rm=TRUE)
```
This is not what we want. It dropped the 80 (i.e. the lowest number and not the NA i.e. missing homework).

Let's look at student3
```{r}
student3
```

```{r}
mean(student3 [ -which.min(student3) ], na.rm=TRUE)
```

One new idea/approach is we could replace the NA (missing homework) with zero.

Let's try with student2
```{r}
student2
```

```{r}
is.na(student2)
```
The **is.na()** function returns a logicalvector where TRUE elements represents where the NA values are.

```{r}
which(is.na(student2))
```
Now let's make the NA values into zeros.

```{r}
student.prime <- student2
student.prime
student.prime[ which(is.na(student.prime)) ] = 0
student.prime
```

Now we need to put this all together to get the average score dropping the lowest where we map NA values to zero.

```{r}
student.prime <- student2
student.prime
student.prime[ which(is.na(student.prime)) ] = 0
mean(student.prime[ -which.min(student.prime) ])
```
```{r}
student.prime
```
Check to see if it works on student3
```{r}
student.prime <- student3
student.prime
student.prime[ which(is.na(student3)) ] = 0
mean(student.prime[ -which.min(student.prime) ])
```
We got out working snippet! Lets simplify. 
```{r}
x <- student3
# Map NA vlaues to zero
x[ which(is.na(x)) ] = 0
# Find the mean without the lowest value
mean(x[ -which.min(x) ])
```


Now we can use this as the body of my function.
```{r}
grade <- function(x) {
#Make sure our scores are all numbers
  x <- as.numeric(x)
# Map NA values to zero
x[ which(is.na(x)) ] = 0
# Find the mean without the lowest value
mean(x[ -which.min(x) ]) }
```

```{r}
grade(student1)
```
```{r}
grade(student2)
```
```{r}
grade(student3)
```

Now read the full gradebook CSV file.
```{r}
scores <- read.csv("https://tinyurl.com/gradeinput", row.names = 1)
scores
```


```{r}
is.numeric(student1)
```
Use for one student
```{r}
grade(as.numeric(scores[2,]))
```
Now grade all students by using the **apply()** function.

```{r}
apply(scores,1,grade)
```

>Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student
overall in the gradebook? [3pts]

```{r}
which.max(apply(scores,1,grade))
```

The top scoring student is student 18.

> Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained
the lowest scores overall? [2pts]

We can now use the apply() function over the columns by setting the margin=2 argument.
```{r}
apply(scores,2,mean,na.rm=TRUE)
```
Homework 3 was the toughest on students.


