---
title: 'class12: Structural Bioinformatics II'
author: "Anel A15426506"
date: "11/4/2021"
output:
  pdf_document: default
  html_document: default
---

\#Comparative analysis of protein structures

using the bio3d package

```{r}
library(bio3d)

pdb <- read.pdb("1hel")
pdb
```

let's use a bioinformarics method called NMA (Normal Mode ANalysis) to predict the dynamics (flexibilty) of this enzyme.

```{r}
modes <- nma(pdb)
plot(modes)
```

make a "move" of its predicted motion. we often call this a "trajectory"

```{r}
mktrj(modes, file="nma.pdb")
```

![](vmdscene.bmp)

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```

# download PDB files

```{r}
#files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Multiple structure alignment

```{r}
#pdb <- pdbaln(files, fit= TRUE)
```

```{r}
pdb
```

```{r}
#ids <- basename.pdb(pdb$id)
```

```{r}
#plot(pdb, labels=ids)
```

we will use the bio3d pca() function which is designed for protein structure data.

```{r}
#pc.xray <- pca(pdb)
#plot(pc.xray)
```

make a trajectory vsiualization of the motion captured by the first Principal component

# Visualize first principal component

```{r}
#pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

![](pc_1.bmp)


