---
title: "class17"
author: "Anel A15426506"
date: "11/28/2021"
output:
  pdf_document: 
    latex_engine: xelatex
---


```{r}
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vacinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

2021-01-05

>Q4. What is the latest date in this dataset? 

2021-11-16

We will use lubridate package to make life a lot easier when dealing with dates and times

```{r}
library(lubridate)
```

```{r}
today()
```

We make our ‘as_of_date’ column lubridate format…

```{r}
# Specify that we
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

## Time difference of 322 days

```{r}
today()- vax$as_of_date[ nrow(vax)]
```


## Time difference of 7 days

Let’s quickly look at the data structure using skim() function

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

8256 missing values
>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

10.17%

>Q8. [Optional]: Why might this data be missing?

>Q9. How many days have passed since the last update of the dataset? 

```{r}
vax$as_of_date[ nrow(vax)] -vax$as_of_date[1]
```
322 days between them

>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
length( unique(vax$as_of_date))
```

47 unique dates

The answer makes sense because

```{r}
47*7
```

We will use **zipcodeR** package to help make sense of the zipcodes

```{r}
library(zipcodeR)
geocode_zip('92037')
```

To calculate distance between two zipcodes:

```{r}
zip_distance('92037','92109')
```

pull census data about ZIP code areas (including median household income etc.):

```{r}
reverse_zipcode(c('92037', "92109"))
```

##Focus on San Diego County

```{r}
table(vax$county)
```

```{r}
inds <- vax$county == "San Diego"
head(vax[inds,])
```

```{r}
library(dplyr)
```

```{r}
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

How many entries are there for San Diego county?

```{r}
nrow(sd)
```

>Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```


>Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
ind <- which.max(sd$age12_plus_population)
sd[ind,]
```

>What is the population in the 92037 ZIP code area?

```{r}
filter(sd, zip_code_tabulation_area == "92037")[1,]
```

>Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 

```{r}
sd.now <- filter(sd, as_of_date == "2021-11-09")
mean(sd.now$percent_of_population_fully_vaccinated, na.rm=TRUE)
```
67.3% are fully vaccinated

>Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

```{r}
library(ggplot2)
ggplot(sd.now) +aes(percent_of_population_fully_vaccinated) + geom_histogram(bins=15)
```

```{r}
ucsd <- filter(sd, zip_code_tabulation_area == "92037")
head(ucsd)
```

```{r}
ucsd[1,]$age5_plus_population
```

>Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)
ggplot(ucsd) +
aes(as_of_date,
percent_of_population_fully_vaccinated) +
geom_point() +
geom_line(group=1) +
ylim(c(0,1)) +
labs(x="Date", y="Percent Vaccinated")
```

##Comparing 92037 to other similar sized areas?

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36.all <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

#head(vax.36)
vax.36 <- filter(vax, age5_plus_population > 36144)
```

```{r}
nrow(vax.36.all)
```

```{r}
length(unique(vax.36.all$zip_code_tabulation_area))
```


>Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
mean(vax.36.all$percent_of_population_fully_vaccinated, na.rm=TRUE)
```

```{r}
ggplot(ucsd) +
aes(as_of_date,
percent_of_population_fully_vaccinated) +
geom_point() +
geom_line(group=1) +
ylim(c(0,1)) +
labs(x="Date", y="Percent Vaccinated") + geom_hline(yintercept=0.66, col="red", linetype="dashed")
```

>Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
summary(vax.36.all$percent_of_population_fully_vaccinated)
```


>Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36.all) +aes(percent_of_population_fully_vaccinated) + geom_histogram(bins=15)
```

>Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%
filter(zip_code_tabulation_area=="92040") %>%
select(percent_of_population_fully_vaccinated)
```

52% less than average
```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%
filter(zip_code_tabulation_area=="92109") %>%
select(percent_of_population_fully_vaccinated)
```
68% above the average




>Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
ggplot(vax.36) + aes(as_of_date,
      percent_of_population_fully_vaccinated,
      group=zip_code_tabulation_area) + geom_line(alpha=0.2) + geom_hline(yintercept = 0.66, col="red", linetype="dashed") + labs(x="Date", y="Percent Vaccinated" )
```
 

>Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I feel a bit uneasy about having class in person because I know most people have traveled and spent their time with big groups.




zz

