---
title: "class09"
author: "Anel A15426506"
date: "10/26/2021"
output: html_document
---

#Data


```{r}
fna.data <- "WisconsinCancer.csv"
```

#completing the code
```{r}
wisc.df <- read.csv(fna.data, row.names = 1)
```

#examine data
```{r}
View(wisc.df)
```

# We can use -1 here to remove the first column
```{r}
wisc.data <- wisc.df[,-1]
```

#Diagnosis vector

```{r}
diagnosis <- as.factor(wisc.df[,1])
```

>Q1. How many observations are in this dataset?

```{r}
nrow(wisc.data)
```
There are 569 observations.


>Q2. How many of the observations have a malignant diagnosis?

```{r}
table(diagnosis)
```
 212 of the observatins have malignant diagnosis.

>Q3. How many variables/features in the data are suffixed with _mean?

```{r}
length(grep("mean", colnames(wisc.df)) )
```
10 features are suffixed with _mean.

# Check column means and standard deviations
```{r}
colMeans(wisc.data)
```


```{r}
apply(wisc.data,2,sd)
```

#PCA on wisc.data
```{r}
wisc.pr <- prcomp(wisc.data, scale. = TRUE)
```

#summary
```{r}
summary(wisc.pr)
```

>Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?

44% of the variance is captured.

>Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?

3 PCs are required to describe at least 70% of the original variance.

>Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

7 PCs are required to describe 90% of the original variance.

#plot
```{r}
biplot(wisc.pr)
```

>Q7. What stands out to you about this plot? Is it easy or difficult to understand? Why?

The plot is not useful because all of the information is mushed together into a blob making it hard to read.

# Scatter plot observations by components 1 and 2

```{r}
plot(wisc.pr$x)
```



```{r}
plot(wisc.pr$x[,1:2], col= factor(diagnosis), 
     xlab = "PC1", ylab = "PC2")
```

>Q8. Generate a similar plot for principal components 1 and 3. What do you notice about these plots?

```{r}
plot(wisc.pr$x[, c(1,3)],col = diagnosis, 
     xlab = "PC1", ylab = "PC3")
```
There isn't a big difference between the graphs besides the data. The data in PC1 VS. PC3 is less negative and more positive.

# Create a data.frame for ggplot
```{r}
df <- as.data.frame(wisc.pr$x)
df$diagnosis <- diagnosis
```


# Load the ggplot2 package
```{r}
library(ggplot2)
```


# Make a scatter plot colored by diagnosis
```{r}
ggplot(df) + 
  aes(PC1, PC2, col=diagnosis) + 
  geom_point()
```

# Calculate variance of each component
```{r}
pr.var <- wisc.pr$sdev^2
head(pr.var)
```

# Variance explained by each principal component: pve
```{r}
pve <- pr.var / sum((pr.var))
```


# Plot variance explained for each principal component
```{r}
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

# Alternative scree plot of the same data, note data driven y-axis
```{r}
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```

>Q9. For the first principal component, what is the component of the loading vector (i.e. wisc.pr$rotation[,1]) for the feature concave.points_mean?

```{r}
wisc.pr$rotation["concave.points_mean",1]
```


>Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?

The minimum number is 8 principal components.

```{r}
summary(wisc.pr)
```

# Scale the wisc.data data using the "scale()" function
```{r}
data.scaled <- ___(wisc.data)
```

