---
title: "class10 halloween"
author: "Anel A15426506"
date: "10/28/2021"
output: html_document
---

```{r}
candy_file <- "candy-data.csv"
```

```{r}
candy = read.csv(url("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv"), row.names=1)
head(candy, n=5)
```

```{r}
View(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy from the list is a Kit Kat.
```{r}
candy["Kit Kat", ]$winpercent
```


>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
```
```{r}
skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

winpercent appears to be on a different scale which is 0-100.


>Q7. What do you think a zero and one represent for the candy$chocolate column?

Zero means no chocolate and 1 means it contains chocolate.

```{r}
skim(candy$chocolate)
```

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```


>Q9. Is the distribution of winpercent values symmetrical?

The values are somewhat symmetrica. However it is more skewed to the left.

>Q10. Is the center of the distribution above or below 50%?

The center is a bit below 50%.

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

Chocolate is higher.

```{r}
inds <- as.logical(candy$chocolate)
chocolate <- candy[inds,]$winpercent

inds.fruit <- as.logical(candy$fruity)
fruity <- candy[inds.fruit,]$winpercent

mean(chocolate)
mean(fruity)
```



>Q12. Is this difference statistically significant?

YES it is significant

```{r}
t.test(chocolate, fruity)
```


>Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```


>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy), winpercent)) +
  geom_col()
```

#add color
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols)
```

Now, for the first time, using this plot we can answer questions like:
>Q17. What is the worst ranked chocolate candy?

The worst ranked is Nik L Nip.
>Q18. What is the best ranked fruity candy?

Starburst


SECTION 4

```{r}
library(ggrepel)
```

```{r}
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese Os Miniatures


```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

The top 5 are shown below and the least popular is Nik L NiP.
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

section 5

```{r}
library(corrplot)
```

```{r}
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity and chocolate
>Q23. Similarly, what two variables are most positively correlated?

chocolate and chocolate


```{r}
pca <- prcomp(candy, scale= TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
```


```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

# Make a new data-frame with our PCA results and candy data
```{r}
my_data <- cbind(candy, pca$x[,1:3])
```
 
 
```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```
 
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```


